---
title: "Project 1"
author: "Calvin Hoang"
date: "1/28/2023"
output:
  pdf_document: default
  html_document: default
---


### Heart Disease Dataset

```{r, echo = FALSE}
heart_disease <- read.csv("C:/Users/hoang/Downloads/STA 106/heart_disease_health_indicators_BRFSS2015.csv", header = TRUE)
```

### Introduction

The Heart Disease Health Indicators Dataset was created from surveys which collect responses from over 400,000 Americans on health-related risk behaviors. In this dataset, I will be analyzing each factor to see which one has effects toward Heart Disease. Furthermore, I will compare which factor has the most and least effects to Heart Disease. However, I will be looking at factors that I believe have certain effects toward Heart Disease.

### Histogram and Analysis

1) Heart Disease vs BMI
```{r, echo = FALSE}
heart_disease$HeartDiseaseorAttack=as.factor(heart_disease$HeartDiseaseorAttack)
```



```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=BMI,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=20,position='stack')+ 
  labs(x ='BMI', title ='Histogram of Heart Disease vs BMI (stack)') 
```


```{r, echo = FALSE}
variable_name='BMI'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 

Based on my observation of this comparison between Heart Disease and BMI, the histogram indicates that the chances of getting heart disease for people who don't have BMI is more likely compared to the people who have BMI. However based on the data chart, we can see that more than half of the odds of each bar indicates the percentage higher than the total odds, while the rest are lower. Furthermore, it shows that the rate of getting heart disease is increasing. The odd ratio between each odds and the odd total is significant. Therefore, BMI has a big effect toward the Heart Disease.

2) Heart Disease vs High Blood Pressure

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=HighBP,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='HighBP', title ='Histogram of High Blood Pressure') 
```



```{r, echo = FALSE}
variable_name='HighBP'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 

- No High Blood Pressure: 23%
- High Blood Pressure: 5%
- There is 4 times difference between these.

For this histogram, it is different from the one above as it only indicates two bars. Based on the histogram, we can see that there is more data of people who don't have high blood pressure. Furthermore, based on the data of the chart, we see that there is more data of people who don't have high blood pressure, and there is a change by 4 times. For high blood pressure, people who have high blood pressure has higher rate of getting heart disease than people who don't have high blood pressure. The odd ratio doesn't seem to be significant.

3) Heart Disease vs High Cholesterol

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=HighChol,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='HighChol', title ='Histogram of Heart Disease vs High Cholesterol') 
```



```{r, echo = FALSE}
variable_name='HighChol'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
- No High Cholesterol: 19%
- High Cholesterol: 5%
- There is 4 times difference between these.

For this histogram, it is similar to the one above. Based on the histogram, we can see that there is more data of people who don't have high cholesterol. Furthermore, based on the data of the chart, we see that there is more data of people who don't have high cholesterol, and there is a change by 4 times. For high cholesterol, people who have high cholesterol has higher rate of getting heart disease than people who don't have high cholesterol. The odd ratio doesn't seem to be significant.

4) Heart Disease vs Cholesterol Check

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=CholCheck,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='CholCheck', title ='Histogram of Heart Disease vs Cholesterol Check') 
```



```{r, echo = FALSE}
variable_name='CholCheck'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
- No Cholesterol Check: 34%
- Cholesterol Check: 9%
- There is a 4 times difference between these.

Based on the histogram, we can see that there is more data of people who have cholesterol check. Furthermore, based on the data of the chart, we see that there is more data of people who have high cholesterol, and there is a change by 4 times. For cholesterol check, people who have cholesterol check has higher rate of getting heart disease than people who don't have cholesterol check. The odd ratio doesn't seem to be significant.

5) Heart Disease vs Smoker

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=Smoker,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='Smoker', title ='Histogram of Heart Disease vs Smoker') 
```



```{r, echo = FALSE}
variable_name='Smoker'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
- No Smoke: 15%
- Smoke: 6%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who don't smoke. Furthermore, based on the data of the chart, we see that there is more data of people who don't have high cholesterol, and there is an change by 2 times. People who smoke has higher rate of getting heart disease than people who don't smoke. The odd ratio doesn't seem to be significant.

6) Heart Disease vs Stroke

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=Stroke,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='Stroke', title ='Histogram of Heart Disease vs Stroke') 
```



```{r, echo = FALSE}
variable_name='Stroke'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 

- No Stroke: 11%
- Stroke: 2%
- There is a 5 times difference between these.

Based on the histogram, we can see that there is more data of people who don't stroke. Furthermore, based on the data of the chart, we see that there is more data of people who don't stroke, and there is a change by 5 times. People who stroke has higher rate of getting heart disease than people who don't stroke. The odd ratio doesn't seem to be significant.



7) Heart Disease vs Physical Activity

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=PhysActivity,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='PhysActivity', title ='Histogram of Heart Disease vs Physical Activity') 
```

```{r, echo = FALSE}
variable_name='PhysActivity'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
- No Physical Activity: 6%
- Physical Activity: 11%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who have physical activity. Furthermore, based on the data of the chart, we see that there is more data of people who have physical activity, and there is a change by 2 times. People who don't have physical activity has higher rate of getting heart disease than people who have. The odd ratio doesn't seem to be significant.

8) Heart Disease vs Fruits

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=Fruits,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='Fruits', title ='Histogram of Heart Disease vs Fruits') 
```

```{r, echo = FALSE}
variable_name='Fruits'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 

- No Fruits: 9%
- Fruits: 10%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who have fruits. Furthermore, based on the data of the chart, we see that there is more data of people who have fruits, and there is a change by 2 times. People who don't have fruits has higher rate of getting heart disease than people who have fruits. The odd ratio doesn't seem to be significant.

9) Heart Disease vs Veggies
```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=Veggies,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='veggies', title ='Histogram of Heart Disease vs Veggies') 
```

```{r, echo = FALSE}
variable_name='Veggies'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 

- No Veggies: 7%
- Veggies: 10%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who have veggies. Furthermore, based on the data of the chart, we see that there is more data of people who have veggies, and there is a change by 2 times. People who don't have veggies has higher rate of getting heart disease than people who have veggies. The odd ratio doesn't seem to be significant.


10) Heart Disease vs Heavy Alcohol Consumption
```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=HvyAlcoholConsump,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='HeavyAlcoholConsumption', title ='Histogram of Heart Disease vs Heavy Alcohol Consumption') 
```

```{r, echo = FALSE}
variable_name='HvyAlcoholConsump'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 

- No Heavy Alcohol Consumption: 9%
- Heavy Alcohol Consumption: 16%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who have heavy alcohol consumption. Furthermore, based on the data of the chart, we see that there is more data of people who have heavy alcohol consumption, and there is a change by 2 times. People who don't have heavy alcohol consumption has higher rate of getting heart disease than people who have veggies. The odd ratio doesn't seem to be significant.

11) Heart Disease vs Any Health Care
```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=AnyHealthcare,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='AnyHealthCare', title ='Histogram of Heart Disease vs Any Health Care') 
```

```{r, echo = FALSE}
variable_name='AnyHealthcare'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
- No Health Care: 13%
- Any Health Care: 9%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who have any health care. Furthermore, based on the data of the chart, we see that there is more data of people who have any health care, and there is a change by 2 times. People who have any health care has higher rate of getting heart disease than people who don't. The odd ratio doesn't seem to be significant.

12) Heart Disease vs No Doctor 
```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=NoDocbcCost,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='NoDocbcCost', title ='Histogram of Heart Disease vs No Doctor because of Cost') 
```

```{r, echo = FALSE}
variable_name='NoDocbcCost'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
- Have Doctor: 10%
- No Doctor: 7%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who have doctors. Furthermore, based on the data of the chart, we see that there is more data of people who have doctors, and there is a change by 2 times. People who don't have a doctor has higher rate of getting heart disease than people who has. The odd ratio doesn't seem to be significant.

13) Heart Disease vs General Health

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=GenHlth,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=20,position='stack')+ 
  labs(x ='GenHlth', title ='Histogram of Heart Disease vs General Health') 
```

```{r, echo = FALSE}
variable_name='GenHlth'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
Based on my observation of this comparison between Heart Disease and General Health, the histogram indicates that the chances of getting heart disease for people who don't have General Health is more likely compared to the people who have General Health. However based on the data chart, we can see that more than half of the odds of each bar indicates the percentage higher than the total odds, while the rest are lower. Furthermore, it shows that the rate of getting heart disease is increasing. The odd ratio between each odds and the odd total seems to be significant. Therefore, General Health has a big effect toward the Heart Disease.

14) Heart Disease vs Mental Health

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=MentHlth,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=20,position='stack')+ 
  labs(x ='MentHlth', title ='Histogram of Heart Disease vs Mental Health') 
```

```{r, echo = FALSE}
variable_name='MentHlth'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
Based on my observation of this comparison between Heart Disease and Mental Health, the histogram indicates that the chances of getting heart disease for people who have Mental Health is more likely compared to the people who don't have Mental Health. Furthermore, based on the data chart, we can see that more than half of the odds of each bar indicates the percentage lower than the total odds. Furthermore, it shows that the rate of getting heart disease is decreasing. The odd ratio between each odds and the odd total is not significant. 
15) Heart Disease vs Physical Health

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=PhysHlth,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=20,position='stack')+ 
  labs(x ='PhysHlth', title ='Histogram of Heart Disease vs Physical Health') 
```

```{r, echo = FALSE}
variable_name='PhysHlth'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
This histogram is aslo similar to the one above. Based on my observation of this comparison between Heart Disease and Physical Health, the histogram indicates that the chances of getting heart disease for people who have Physical Health is more likely compared to the people who don't have Physical Health. Furthermore, based on the data chart, we can see that more than half of the odds of each bar indicates the percentage lower than the total odds. Furthermore, it shows that the rate of getting heart disease is decreasing. The odd ratio between each odds and the odd total is not significant. 


16) Heart Disease vs Sex

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=Sex,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=3,position='stack')+ 
  labs(x ='Sex', title ='Histogram of Sex') 
```

```{r, echo = FALSE}
variable_name='Sex'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
- No Health Care: 13%
- Any Health Care: 7%
- There is a 2 times difference between these.

Based on the histogram, we can see that there is more data of people who don't have sex. Furthermore, based on the data of the chart, we see that there is more data of people who don't have sex, and there is a change by 2 times. People who have sex has higher rate of getting heart disease than people who don't. The odd ratio doesn't seem to be significant.


17) Heart Disease vs Income

```{r, echo = FALSE}
library(ggplot2)
ggplot(data=heart_disease, aes(x=Income,fill = HeartDiseaseorAttack)) + 
  geom_histogram(bins=20,position='stack')+ 
  labs(x ='Income', title ='Histogram of Heart Disease vs Income') 
```


```{r, echo = FALSE}
variable_name='Income'
response_name='HeartDiseaseorAttack'

ax=seq(min(heart_disease[,variable_name]),max(heart_disease[,variable_name]),length.out=10)
ax
hgA = hist(heart_disease[heart_disease[,response_name]==0,variable_name], breaks = ax, plot = FALSE)
hgB = hist(heart_disease[heart_disease[,response_name]==1,variable_name], breaks = ax, plot = FALSE)

table1=rbind('FALSE'=hgA$counts,'TRUE'=hgB$counts)

colnames(table1)=round(hgA$breaks[-1],0)
# add row sum
table1=cbind(table1, 'Total'=apply(table1,1,sum))
# add column sum
table1=rbind(table1, 'Total'=apply(table1,2,sum))
# add odds
table1=rbind(table1, 'Odds'=round(table1[1,]/table1[2,],2))
to_print='Odds'=round(table1[1,]/table1[2,],2)
to_print_ratio='ratio'=round(table1[2,]/table1[1,],2)

table1
``` 
Based on my observation of this comparison between Heart Disease and Income, the histogram indicates that the chances of getting heart disease for people who don't have Income is more likely compared to the people who have Income. Furthermore, based on the data chart, we can see that more than half of the odds of each bar indicates the percentage lower than the total odds. However, it shows that the rate of getting heart disease is increasing. The odd ratio between each odds and the odd total is not significant. 

### Conclusion

After going through comparing each factors to the heart disease, I conclude that only the BMI and General Health have significant effects to the Heart Disease since they have significant odd ratio. Furthermore, although the rest of the factors don't have significant effects to the Heart Disease, it can be seen that they still have effects at a certain degree.

### Code Appendix
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}

```